Abstractions for Multi-Sorted Substitutions

### 2023-09-16 paper/paper.tex
Update.

  \section{Introduction}
  In recent years intrinsic typing\cite{TODO} has become rather popular.
  A consequence of intrinsic typing is that operations on terms
  are also proofs that the operations preserve typing.
  This often has the benefit of more concise code, as shared structure
  between definitions and preservation lemmas does not need to be
  repeated.
  However, intrinsic typing also has three major drawbacks:
  \begin{enumerate}
  \item
    It is more difficult to judge whether a specification is correct,
    because there is no clear separation between specification and proofs.
    To understand the specification one has to additionally
    understand the proofs, which can introduce significant cognitive overhead.
    % For example specifying a small step semantics is directly intertwined
    % with the proof of subject reduction.
  \item
    It is less uniform for object languages with multiple variable
    sorts, e.g. type- and expression-variables. In such a scenario,
    types are usually indexed by kinds and expressions by kinds and types.
    This requires not only a type-in-type and expression-in-expression
    substitution, but also a type-in-expression substitution, which causes
    interactions between different types of substitutions introducing
    additional complexity.
    The duplication increases even more if one additionally introduces
    kind-variables. Then a total of 6 different substitutions and
    their interactions is required.
    Unifying these substitutions to reduce repetition is difficult
    with intrinsic typing due to the interdependence of definitions.
  \item
    It is less suitable for automation, because type preservation is baked
    into the syntax definition and type preservation can become arbitrary complex
    depending on the typing relation. This can prevent substitition
    definition and lemmas from being derived, where otherwise it would be possible.
  \end{enumerate}
  With extrinsic typing the major drawback is the duplication between
  definitions and type preservation lemmas, which we found more
  suitable to be eliminated by automation.

@@

  \subsection{Structure}
  The rest of this paper will introduce our framework using System F
  as a running example for a type soundness proof.

  Code of the framework is displayed in gray boxes.
  Code of examples is written in yellow boxes.
  Code of the System F formalization is written without boxes. This is
  the only code a user of our framework has to write.

  In the paper we state all definitions and lemmas, but omit some proofs
  in the interest of space. However the omitted proofs exclusively
  depend on lemmas and definitions stated in this paper. The full code
  from the paper can be found at TODO.

  In the paper we only show the part of the framework necessary for
  the System F soundness proof. The actual implementation is more
  general and complete and can be found at TODO.

  \subsection{Contributions}
  \begin{enumerate}
  \item
    a multi-sorted approach to extrinsic typing;
  \item
    a multi-sorted extension of McBride's kits to unify renamings and substitutions and their metatheory;
  \item
    an abstraction for composition and it's metatheory, unifying the four combinations of renamings and substitutions;
  \item
    an abstraction for typing relations, unifying type preservation of renaming and substitution into a single lemma;
  \item
    an Agda-implementation featuring derived term-kits, representation
    independence for substitutions and typing contexts, heterogeneous
    equality, independence of functional extensionality, case studies
    on System F, dependent lambda calculus, pattern matching, linear
    types.
  \end{enumerate}

  \newpage

  \section{Syntax}
  \subsection{Single-Sorted Syntax}
  The following shows a typical intrinsically-scoped syntax of System F:
  \ExampleCode\FUnsortedSyntax

  \texttt{Type}s are indexed by the number of free type variables \texttt{n}.
  \texttt{Expr}essions are additionally indexed by the number of free expression variables \texttt{m}.
  Variables \texttt{`\_} are represented as DeBruijn-indices, where \texttt{Fin n} is the type of \texttt{n} elements.

  We identify two drawbacks with this style of syntax:
  \begin{enumerate}
  \item the syntactic categories (\texttt{Kind}, \texttt{Type}, and \texttt{Expr})
    have different types, which makes it difficult to treat them uniformly; and
  \item the different sorts of variables are modeled separately, which requires to define not just
    type-in-type and expression-in-expression substitution, but also type-in-expression substitution.
    Consequently, interaction lemmas between those substitutions are required.
    In general, this leads to a combinatory explosion of
    substitutions, e.g.\ adding kind variables would lead to a total of 6
    different substitutions.
  \end{enumerate}
  To avoid these drawbacks, we instead use a multi-sorted syntax.

  \subsection{Multi-Sorted Syntax}
  A multi-sorted syntax is defined by a single type of sort-indexed terms.

@@

  The term type \ACode{S \ACon{⊢} s} is then indexed by its sort
  \ACode{s} and the sorts of its free variables \ACode{S}.
  For example, \ACode{\ACon{[𝕥, 𝕥] ⊢ 𝕖}} is the type of expressions
  (\ACode{\ACon{𝕖}}) with two free type-variables (\ACode{\ACon{𝕥}}).
  \newpage
  \AppCode\FSyntax
  The notation \ACode{\ACon{\_⊢\_}} is often used for terms in
  intrinsically-typed languages. This is no accident: in effect, we
  defined an intrinsically-typed language with the twist that the
  typing relation assures exactly that the syntactic categories are followed.
  Sorts \ACode{s} correspond to types, and lists of sorts \ACode{S}
  correspond to type environments.

@@

  \subsection{A Structure for Multi-Sorted Syntax}
  The multi-sorted syntax makes it easy to define a structure for
  syntaxes of arbitrary object-languages, i.e.\ syntaxes with
  arbitrary amounts of syntactic categories and variables:
  \LibCode\KSyntax
  The first three fields record the definitions of sorts, terms, and variable introduction.
  The last field records that variable introduction
  \ACode{\AField{`\_}} is injective, which is trivially true for
  constructors. The instantiation for our System F syntax is
  straightforward:
  \AppCode\FSyntaxInst

  \section{Renamings \& Substitutions}
  \subsection{Multi-Sorted Renamings \& Substitutions}
  Working with a sort-indexed syntax allows us to define renamings and substitutions
  that replace all variables of all sorts simultaneously:

@@

  \subsection{The Need for Renamings}
  A problem with mechanizing languages with DeBruijn-indices is that
  the operation of applying a substitution \ACode{σ} to a term
  \ACode{t} cannot be defined directly by structural recursion on
  \ACode{t}.
  If \ACode{t} is a term like \ACode{\ACon{λx} e},
  where the subterm \ACode{e} is under a binder, then the variables
  in \ACode{e} are shifted, i.e.\ the variable \ACode{\ACon{zero}} refers
  to the variable from the binder, whereas a variable \ACode{\ACon{suc} x}
  refers to the variable \ACode{x} from outside of the binder.
  As the terms contained in \ACode{σ} may themselves contain free variables,
  they need to be shifted, too, before they can be substituted
  for variables in \ACode{e}. 
  Shifting the variables in a term is itself a substitution, which maps
  each variable \ACode{x} to the term \ACode{\ACon{`} (\ACon{suc} x)}.
  However, applying this shifting substitution to the terms in
  \ACode{σ} is not structurally recursive, as the terms in \ACode{σ}
  are not subterms of \ACode{\ACon{λx} e}.

@@

  \subsection{Unifying Renamings \& Substitutions}
  To avoid the duplication between renamings and substitutions,
  McBride\cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}
  introduced an abstraction called \emph{kits}.\footnote{
    While kits were originally formulated for intrinsically-typed languages,
    they nicely extend to our extrinsically-typed multi-sorted syntax, whose
    sort-indexing is a form of intrinsic typing.
  }

@@

  \section{Map Composition}

@@

  \subsection{An Examination of Composition}
  To motivate our abstraction for composition, we first look
  at the four compositions individually:
  \ExampleCode\KExFourComps
  The definitions reveal two interesting properties:
  \begin{enumerate}
  \item
    If we compose two maps \ACode{ϕ₁} and \ACode{ϕ₂}, then the
    resulting map is a renaming, iff both \ACode{ϕ₁} and \ACode{ϕ₂} are
    renamings.
    In other words, 
    if \ACode{ϕ₁} is a \ACode{K₁}-map and \ACode{ϕ₂} is a \ACode{K₂}-map,
    then the result is a \ACode{(K₁ ⊔ K₂)}-map, where \ACode{⊔} refers to the
    lattice for \ACode{\{ \ADef{Kᵣ} , \ADef{Kₛ} \}}
    generated by \ACode{\ADef{Kᵣ} $\leq$ \ADef{Kₛ}}.
  \item
    All four compositions first apply \ACode{ϕ₁} to \ACode{x},
    and then apply \ACode{ϕ₂} to the result.
    If \ACode{ϕ₁} is a renaming, this result is another variable, but
    if \ACode{ϕ₁} is a substitution, this result is a term.
  \end{enumerate}
  With the \ACode{\ACon{Kit}} abstraction, we can easily abstract over
  \ACode{ϕ₂} being a renaming or a substitution:
  \ExampleCode\KExTwoComps

@@

  \subsection{WkKits}

@@

  \subsection{Composition}

@@

  \section{Types \& Contexts}
  % To define typing contexts and a typing relation, we need to know
  % which sorts of terms play the role of types for other sorts of terms.
  % When modeling a typed language with multi-sorted syntax, both terms
  % and types are represented as terms, but with different sorts.
  % To define typing contexts and a typing relation, we need to record for each term
  % from 
  % describe terms, and ocan be mapped to 
  % In a typed language,

@@

  \section{Typing}
  Equipped with a notion of types and type context we are ready to define
  the multi-sorted typing relation for System F, incorporating both
  typing and kinding:
  \AppCode\FTyping
  The interesting cases are:
  \begin{itemize}
  \item
    the variable typing constructor covers both expression- and
    type-variables, analogously to the variable term constructor;
  \item
    the typing of the lambda abstraction \ACode{\ACon{⊢λ}} weakens the
    type of the codomain \ACode{t₂}.
    This is necessary, because with multi-sorted syntax, types could
    also depend on expressions, so the sub typing derivation accounts for
    another variable which is not used by the type.
  \item
    the kinding constructor \ACode{\ACon{⊢τ}} states that all types
    have kind \ACode{\ACon{★}}. This is sufficient for System F as types
    are automatically well-scoped due to intrinsic scoping.
  \end{itemize}
  To teach the framework about typing, we create a structure analogously to
  \ACode{\ACon{Syntax}}:
  \LibCode\KTyping
  and instanciate it for System F:
  \AppCode\FTypingInst

@@

  \section{Semantics}
  We define a standard small-step operational semantics for System F
  with full reduction:
  \AppCode\FReduction

  \section{Subject Reduction}
  Proving subject reduction is now straightforward:
  \AppCode\FSubjectReduction
  \AppCode\FSubjectReductionProofInteresting
  We only listed the cases that do not follow directly from the
  induction hypothesis:
  \begin{itemize}
  \item The beta-reduction rule for type-application, \ACode{\ACon{β-Λ}},
    is a direct consequence of substitution preserves typing.
    \ACode{⊢e₂} is the typing for \ACode{e₂}.
    \ACode{⊢⦅ ⊢e₂ ⦆} is the substitution-typing for the singleton substitution 
    \ACode{\AField{⦅ e₂ ⦆}}.
    \ACode{\AField{⊢e₁ ⊢⋯ ⊢⦅ ⊢e₂ ⦆}} states that applying the singleton substitution
    to \ACode{e₁} preserves typing.
  \item The beta-reduction rule for value-application is almost the same,
    but requires us to addionaly apply an equality to the type index.
  \end{itemize}

@@

  \section{Reflection \& Generics}
  We used Agda's reflection mechanism\cite{TODO} to derive
  instanciations related to all structures for untyped substitition,
  i.e. \ACode{\ACon{Traversal}} and \ACode{\ACon{ComposeTraversal}}.

@@

  \section{Beyond System F}
  \begin{itemize}
  \item
    Binders with multiple variables are supported and even pattern
    matching can be encoded.
  \item
    Heterogenous equality
  \end{itemize}

  \section{Future Work}
  \begin{itemize}
  \item
    Enabling more object languages by allowing subterms with less
    restrictions on the form of \ACode{S}.
  \end{itemize}

  \section{Related Work}
  Autosubst \cite{DBLP:conf/cpp/StarkSK19, DBLP:conf/itp/SchaferTS15}.
  Kits \cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}.
  Kits linear \cite{DBLP:journals/corr/abs-2005-02247}.
  Universe \cite{DBLP:journals/pacmpl/AllaisA0MM18}.

  \section{Conclusion}

>>>>

  \section{Introduction}
  \label{sec:introduction}
  In recent years intrinsic typing\cite{TODO} has become rather popular.
  A consequence of intrinsic typing is that operations on terms
  are also proofs that the operations preserve typing.
  This often has the benefit of more concise code, as shared structure
  between definitions and preservation lemmas does not need to be
  repeated.
  However, intrinsic typing also has three major drawbacks:
  \begin{enumerate}
  \item
    It is more difficult to judge whether a specification is correct,
    because there is no clear separation between specification and proofs.
    To understand the specification one has to additionally
    understand the proofs, which can introduce significant cognitive overhead.
    % For example specifying a small step semantics is directly intertwined
    % with the proof of subject reduction.
  \item
    It is less uniform for object languages with multiple variable
    sorts, e.g. type- and expression-variables. In such a scenario,
    types are usually indexed by kinds and expressions by kinds and types.
    This requires not only a type-in-type and expression-in-expression
    substitution, but also a type-in-expression substitution, which causes
    interactions between different types of substitutions introducing
    additional complexity.
    The duplication increases even more if one additionally introduces
    kind-variables. Then a total of 6 different substitutions and
    their interactions is required.
    Unifying these substitutions to reduce repetition is difficult
    with intrinsic typing due to the interdependence of definitions.
  \item
    It is less suitable for automation, because type preservation is baked
    into the syntax definition and type preservation can become arbitrary complex
    depending on the typing relation. This can prevent substitition
    definition and lemmas from being derived, where otherwise it would be possible.
  \end{enumerate}
  With extrinsic typing the major drawback is the duplication between
  definitions and type preservation lemmas, which we found more
  suitable to be eliminated by automation.

@@

  \subsection{Structure}
  \label{sec:introduction:structure}
  The rest of this paper will introduce our framework using System F
  as a running example for a type soundness proof.

  Code of the framework is displayed in gray boxes.
  Code of examples is written in yellow boxes.
  Code of the System F formalization is written without boxes. The
  latter is the only code a user of our framework has to write.

  In this paper we present a simplified version of the actual
  framework, focusing on the core concepts. We present all necessary
  definitions and lemmas, but omit some proofs for the sake of
  space. The omitted proofs can be found in Appendix~{TODO} and the
  supplementary material. The full framework can be found at TODO.

  The rest of this paper is structured as followed:
  Section~\ref{sec:syntax} introduces the multi-sorted syntax and compares it to the
  more common unsorted syntax.
  Section~\ref{sec:maps} introduces multi-sorted substitutions and renamings, and
  an abstraction to unify them.
  Section~\ref{sec:composition} introduces composition of multi-sorted
  substitions and renamings, and an abstraction to unify all four
  compositions.
  Section~\ref{sec:types} shows how to introduce the notion of types
  for multi-sorted syntax and how to define typing contexts.
  Section~\ref{sec:typing} shows how to define a multi-sorted typing relation,
  and an abstraction to unify type preservattion lemmas for renamings and substitutions.
  Section~\ref{sec:semantics} introduces a small-step operational semantics for System F.
  Section~\ref{sec:subjectreduction} proves the subject reduction lemma for System F.
  Section~\ref{sec:reflection} describes the class of object languages covered by
  our reflection algorithm.
  Section~\ref{sec:beyond} describes case studies of object languages beyond System F.
  Section~\ref{sec:future} describes future work.
  Section~\ref{sec:related} discusses related work.
  Section~\ref{sec:conclusion} concludes.

  \subsection{Contributions}
  \label{sec:introduction:contributions}
  \begin{enumerate}
  \item
    a novel approach for formalizing intrinsically-scoped syntax with
    multiple variable sorts as a special kind of intrinsically-typed
    syntax, we call \emph{multi-sorted syntax};
  \item
    an adaptation of McBride's kit-abstraction for unifying renamings
    and substitutions to multi-sorted syntax;
  \item
    a novel abstraction for composition and it's metatheory, unifying
    the four compositions between renamings and substitutions;
  \item
    a novel abstraction for typing relations, unifying type
    preservation of renaming and substitution;
  \item
    a formalized specification of a large class of object languages
    for which substitution and lemmas can be derived generically.
  \item
    an implementation of the concepts as framework in Agda featuring
    a reflection algorithm, representation independence for
    substitutions and typing contexts, heterogeneous equality between
    renamings and substitutions, and free of postulated axioms.
  \item
    three case studies using our approach to prove subject reduction
    for System~F with subtyping, a dependently-typed lambda calculus, and
    pattern matching.
  \end{enumerate}

  \section{Syntax}
  \label{sec:syntax}
  \subsection{Single-Sorted Syntax}
  \label{sec:syntax:single}
  The following shows a typical intrinsically-scoped syntax of System F:
  \ExampleCode\FUnsortedSyntax

  \ACode{\ACon{Type}}s are indexed by the number of free type
  variables \ACode{n}.
  \ACode{\ACon{Expr}}essions are additionally indexed by the number of
  free expression variables \ACode{m}.
  Variables \ACode{\ACon{`\_}} are represented as DeBruijn-indices, where
  \ACode{\ACon{Fin} n} is the type of \ACode{n} elements.

  We identify two drawbacks with this style of syntax:
  \begin{enumerate}
  \item
    the syntactic categories (\ACode{\ACon{Kind}},
    \ACode{\ACon{Type}}, and \ACode{\ACon{Expr}})
    have different types, which makes it difficult to treat them uniformly; and
  \item the different sorts of variables are modeled separately, which requires to define not just
    type-in-type and expression-in-expression substitution, but also type-in-expression substitution.
    Consequently, interaction lemmas between those substitutions are required.
    In general, this leads to a combinatory explosion of
    substitutions, e.g.\ adding kind variables would lead to a total of 6
    different substitutions and corresponding interaction lemmas.
  \end{enumerate}
  To avoid these drawbacks, we instead use a multi-sorted syntax.

  \subsection{Multi-Sorted Syntax}
  \label{sec:syntax:multi}
  A multi-sorted syntax is defined by a single type of sort-indexed terms.

@@

  The term type \ACode{S \ACon{⊢} s} is then indexed by its sort
  \ACode{s} and the sorts of its free variables \ACode{S}.
  For example, \ACode{\ACon{[𝕥, 𝕥] ⊢ 𝕖}} is the type of expressions
  (\ACode{\ACon{𝕖}}) with two free type-variables (\ACode{\ACon{𝕥}}).
  \AppCode\FSyntax
  The notation \ACode{\ACon{\_⊢\_}} is often used for terms in
  intrinsically-typed languages. This is no accident: in effect, we
  defined an intrinsically-typed language with the twist that the
  typing relation assures exactly that the syntactic categories are followed.
  Sorts \ACode{s} correspond to types, and lists of sorts \ACode{S}
  correspond to type environments.

@@

  \subsection{A Structure for Multi-Sorted Syntax}
  \label{sec:syntax:structure}
  The multi-sorted syntax makes it easy to define a structure for
  syntaxes of arbitrary object-languages, i.e.\ syntaxes with
  arbitrary amounts of syntactic categories and variables:
  \LibCode\KSyntax
  The first three fields record the definitions of sorts, terms, and variable introduction.
  The last field records that variable introduction
  \ACode{\AField{`\_}} is injective, which is trivially true for
  constructors. The instantiation for our System F syntax is
  straightforward:
  \AppCode\FSyntaxInst

  \section{Renamings \& Substitutions}
  \label{sec:maps}
  \subsection{Multi-Sorted Renamings \& Substitutions}
  \label{sec:maps:example}
  Working with a sort-indexed syntax allows us to define renamings and substitutions
  that replace all variables of all sorts simultaneously:

@@

  \subsection{The Need for Renamings}
  \label{sec:maps:renamings}
  A problem with mechanizing languages with DeBruijn-indices is that
  the operation of applying a substitution \ACode{σ} to a term
  \ACode{t} cannot be defined directly by structural recursion on
  \ACode{t}.
  If \ACode{t} is a term like \ACode{\ACon{λx} e},
  where the subterm \ACode{e} is under a binder, then the variables
  in \ACode{e} are shifted, i.e.\ the variable \ACode{\ACon{zero}} refers
  to the variable from the binder, whereas a variable \ACode{\ACon{suc} x}
  refers to the variable \ACode{x} from outside of the binder.
  As the terms contained in \ACode{σ} may themselves contain free variables,
  they need to be shifted, too, before they can be substituted
  for variables in \ACode{e}. 
  Shifting the variables in a term is itself a substitution, which maps
  each variable \ACode{x} to the term \ACode{\ACon{`} (\ACon{suc} x)}.
  However, applying this shifting substitution to the terms in
  \ACode{σ} is not structurally recursive, as the terms in \ACode{σ}
  are not subterms of \ACode{\ACon{λx} e}.

@@

  \subsection{Unifying Renamings \& Substitutions}
  \label{sec:maps:kits}
  To avoid the duplication between renamings and substitutions,
  McBride\cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}
  introduced an abstraction called \emph{kits}.\footnote{
    While kits were originally formulated for intrinsically-typed languages,
    they nicely extend to our extrinsically-typed multi-sorted syntax, whose
    sort-indexing is a form of intrinsic typing.
  }

@@

  \section{Map Composition}
  \label{sec:composition}

@@

  \subsection{An Examination of Composition}
  \label{sec:composition:example}
  To motivate our abstraction for composition, we first look
  at the four compositions individually:
  \ExampleCode\KExFourComps
  The definitions reveal two interesting properties:
  \begin{enumerate}
  \item
    If we compose two maps \ACode{ϕ₁} and \ACode{ϕ₂}, then the
    resulting map is a renaming, iff both \ACode{ϕ₁} and \ACode{ϕ₂} are
    renamings.
    In other words, 
    if \ACode{ϕ₁} is a \ACode{K₁}-map and \ACode{ϕ₂} is a \ACode{K₂}-map,
    then the result is a \ACode{(K₁ ⊔ K₂)}-map, where \ACode{⊔} refers to the
    lattice for \ACode{\{ \ADef{Kᵣ} , \ADef{Kₛ} \}}
    generated by \ACode{\ADef{Kᵣ} $\leq$ \ADef{Kₛ}}.
  \item
    All four compositions first apply \ACode{ϕ₁} to \ACode{x},
    and then apply \ACode{ϕ₂} to the result.
    If \ACode{ϕ₁} is a renaming, this result is another variable, but
    if \ACode{ϕ₁} is a substitution, this result is a term.
  \end{enumerate}
  With the \ACode{\ACon{Kit}} abstraction, we can easily abstract over
  \ACode{ϕ₂} being a renaming or a substitution:
  \ExampleCode\KExTwoComps

@@

  \subsection{WkKits}
  \label{sec:composition:wkkit}

@@

  \subsection{Composition}
  \label{sec:composition:ckit}

@@

  \section{Types \& Contexts}
  \label{sec:types}
  % To define typing contexts and a typing relation, we need to know
  % which sorts of terms play the role of types for other sorts of terms.
  % When modeling a typed language with multi-sorted syntax, both terms
  % and types are represented as terms, but with different sorts.
  % To define typing contexts and a typing relation, we need to record for each term
  % from 
  % describe terms, and ocan be mapped to 
  % In a typed language,

@@

  \section{Typing}
  \label{sec:typing}
  Equipped with a notion of types and type context we are ready to define
  the multi-sorted typing relation for System F, incorporating both
  typing and kinding:
  \AppCode\FTyping
  The interesting cases are:
  \begin{itemize}
  \item
    the variable typing constructor covers both expression- and
    type-variables, analogously to the variable term constructor;
  \item
    the typing of the lambda abstraction \ACode{\ACon{⊢λ}} weakens the
    type of the codomain \ACode{t₂}.
    This is necessary, because with multi-sorted syntax, types could
    also depend on expressions, so the sub typing derivation accounts for
    another variable which is not used by the type.
  \item
    the kinding constructor \ACode{\ACon{⊢τ}} states that all types
    have kind \ACode{\ACon{★}}. This is sufficient for System F as types
    are automatically well-scoped due to intrinsic scoping.
  \end{itemize}
  To teach the framework about typing, we create a structure analogously to
  \ACode{\ACon{Syntax}}:
  \LibCode\KTyping
  and instanciate it for System F:
  \AppCode\FTypingInst

@@

  As the \ACode{\ACon{Traversal}} and  \ACode{\ACon{ComposeTraversal}} instances
  can be generally derived via reflection,
  the \ACode{\AField{\_⊢⋯\_}} lemma is in fact the only lemma that
  has to been proved manually for subject reduction.
  From this single lemma follows type preservation for all operations
  of both renamings and substitutions for all sorts of variables
  simultaneously.

  \section{Semantics}
  \label{sec:semantics}
  We define a standard small-step operational semantics for System F
  with full reduction:
  \AppCode\FReduction

  \section{Subject Reduction}
  \label{sec:subjectreduction}
  Proving subject reduction is now straightforward:
  \AppCode\FSubjectReduction
  \AppCode\FSubjectReductionProofInteresting
  We only listed the cases that do not follow directly from the
  induction hypothesis:
  \begin{itemize}
  \item The beta-reduction rule for type-application, \ACode{\ACon{β-Λ}},
    is a direct consequence of substitution preserves typing.
    \ACode{⊢e₂} is the typing for \ACode{e₂}.
    \ACode{⊢⦅ ⊢e₂ ⦆} is the substitution-typing for the singleton substitution 
    \ACode{\AField{⦅ e₂ ⦆}}.
    \ACode{\AField{⊢e₁ ⊢⋯ ⊢⦅ ⊢e₂ ⦆}} states that applying the singleton substitution
    to \ACode{e₁} preserves typing.
  \item The beta-reduction rule for value-application is almost the same,
    but requires us to addionaly apply an equality to the type index.
  \end{itemize}

@@

  \section{Reflection \& Generics}
  \label{sec:reflection}
  We used Agda's reflection mechanism\cite{TODO} to derive
  instanciations related to all structures for untyped substitition,
  i.e. \ACode{\ACon{Traversal}} and \ACode{\ACon{ComposeTraversal}}.

@@

  \section{Beyond System F}
  \label{sec:beyond}
  \begin{itemize}
  \item
    Binders with multiple variables are supported and even pattern
    matching can be encoded.
  \item
    Heterogenous equality
  \end{itemize}

  \section{Future Work}
  \label{sec:future}
  \begin{itemize}
  \item
    Enabling more object languages by allowing subterms with less
    restrictions on the form of \ACode{S}.
  \item
    Providing abstractions not just for type-preservation, but also
    for different kinds of semantics, e.g. confluence or lemmas related
    to denotational semantics.
  \end{itemize}

  \section{Related Work}
  \label{sec:related}
  Autosubst \cite{DBLP:conf/cpp/StarkSK19, DBLP:conf/itp/SchaferTS15}.
  Kits \cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}.
  Kits linear \cite{DBLP:journals/corr/abs-2005-02247}.
  Universe \cite{DBLP:journals/pacmpl/AllaisA0MM18}.

  \section{Conclusion}
  \label{sec:conclusion}
